Use of general purpose graphical processing units with MODFLOW

### 2012-08-31 ResponseTo_GW20120518-0099_Review.tex

\item \textit{Line 47 "present" (plural)}

Correction made.

\item \textit{Line 48 - unclear - faster at what exactly? Operations,cycles/second? }

This sentence has been revised to indicate that GPGPUs are capable of executing more floating point operations per second than CPUs.

@@

\item \textit{Line 108 - really, the parallelization you mention can be applied to many simulators, not just MODFLOW-2005. Consider mentioning wider range applicability. }

We have added text to indicate that GPGPU parallelization has great potential for MODFLOW and other numerical simulators.

\item \textit{Line 213-214 - unclear. Consider rephrasing }

This sentence has been rephrased to clarify the description of the thickness of layer 1.

@@

\item \textit{Line 351 - "developed to *use* the PCG" (one does not "solve" the algorithm) }

Suggested modification made.

\item \textit{Line 376 - as indicated above, can you provide some guidance here about when one approach might be better than another depending on the system properties (size, K field). Also, GMG should be included here to so that the full suite of approaches are compared (weak but highly parallelizable to strong but not very parallelizable). }

@@

\section*{Reviewer 3 Comments:}
\begin{enumerate}
\item \textit{Quality of figures should be improved (all figures but in particular, figures 3 to 6). Texts (titles and legends) are too small to read. Lines are very hard to read.}

The font sizes have been increased on all of the figures. The symbol size and lines on \textbf{figures 3}, \textbf{5}, and \textbf{6} have been increased and converted to solid colored lines, respectively, to improve the readability as suggested.

\item \textit{Tables should be reworked. I do not think we need 3 digits after decimal point for these comparisons. Please also consider significant digits in the tables.}

Double-precision floating point variables are used in the solvers and the tables have been rounded to single-precision numbers. Since we are reporting results from a numerical analysis we thought it appropriate to retain a larger number of digits to allow readers the ability to evaluate our results differently than we have. For example it would be possible for readers to calculate accurate run times (in seconds) for each of the simulations summarized in \textbf{table 2} using the speedup values and MIC cpu times. This may be of benefit for future studies conducted by other researchers. However, we would be willing to modify the number of digits reported if the editor agrees with this comment.

@@

\item \textit{The authors only consider total execution time which includes time required for both Picard and linear solver iterations. Did you closely examined these iteration performance in solver comparisons? How these iteration performance changes among different preconditioning schemes? }

\textbf{Table 4} has been added which includes information on the number of iterations required for each of the solvers and various preconditioners. Additional discussion has been added in the \textbf{Discussion} section to provide some guidance on the strength of the preconditioners for the problems evaluated.

@@

\item \textit{Lines 28-32: the authors mentioned computational performance changes due to domain discretization. Does this cause by FD formulation or matrix solver scheme? Explanation is needed. }

This is primarily based on the increase in the number of simultaneous linear equations being solved. Differences in the \textbf{A} matrix resulting from different cell sizes can also affect solver performance. Additional clarification has been added.

@@

\item \textit{Line 161: What is CUBLAS? }

CUBLAS and BLAS have been defined.

\item \textit{Line 199: Problem Description. Reasoning for choosing test cases is helpful for the readers. }

Additional discussion of the reasons the particular test cases were evaluated has been provided.

### 2012-08-31 GW20120518-0099-r01.tex
Additional edits

%%% PAGE DIMENSIONS
\usepackage{geometry}
\usepackage{lineno}
\usepackage{color}
\usepackage[kernelfbox]{xcolor}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
%%% LANGUAGE HYPHENATION
\usepackage[english]{babel}
%%% HEADERS & FOOTERS
\usepackage{fancyhdr} 
%%% SECTION TITLE APPEARANCE
\usepackage{sectsty}
%%% MATHS RELATED PACKAGES
\usepackage{import}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{setspace}
\usepackage{array}
\usepackage{multirow}
%%% OTHER PACKAGES
\usepackage{booktabs}
\usepackage{array}
\usepackage{paralist}
\usepackage{verbatim}
\usepackage{subfig}
\usepackage{url}

@@

\color{black}Dong and Li (2009) parallelized the MODFLOW PCG solver (Hill 1990) using the OpenMP (OpenMP Architecture Review Board, 2005) programming paradigm and observed speedups as large as 5 on machines with multi-core CPUs. Naff (2008) developed a parallelized linear solver for MODFLOW using the Message Passing Interface (MPI) standard and observed a speedup of 7. In addition to OpenMP and MPI, GPGPUs also \color{blue}present \color{black}an option for linear solver parallelization. In comparison to CPUs, GPGPUs are much faster, have higher bandwidth, and typically more cores (\textit{e.g.}, 448 \color{blue} cores \color{black} in the case of the NVIDIA$^{\textregistered}$ Tesla™ C2050 and C2070 GPGPUs). Use of GPGPUs to parallelize linear solvers for MODFLOW is also attractive because of the availability of the CUDA (Compute Unified Device Architecture) application programming interface (API) for NVIDIA GPGPUs, which supports development of parallel GPGPU code in C/C++ or Python, Fortran, Java, and MATLAB using wrappers around the C/C++ implementation. For example, Soni \textit{et al.} (2012) used the CUDA language to develop a GPGPU code to solve computational fluid dynamics problems. For solution of the groundwater flow equation, Ji \textit{et al.} (2012) developed a parallel GPGPU version of MODFLOW-2000 with a Jacobi preconditioner and observed speedups ranging from 1.8 to 3.7.

@@

For GPGPU code implementations, the concept is similar to OpenMP.  The main code is a sequential host code which forks the data into many threads at the invocation of a CUDA \color{blue}functions, called kernels (\textbf{Box 1})\color{black}.  \color{blue}However, as opposed to the few threads spawned by OpenMP, GPGPUs have to potential to spawn thousands to tens-of-thousands of threads to execute the CUDA kernel function (\textbf{Box 2}). The NVIDIA Tesla C2050 used to evaluate the UPCG solver in this study can execute 21,504 concurrent threads. \color{black} With thread counts on the order of tens of thousands, GPGPUs have great potential for parallelization \color{blue} not just for MODFLOW, but for most numerical modeling applications\color{black}.

@@

\subsection*{Problem Description}
GPGPU and parallel CPU results were evaluated using a number of horizontal and vertical model discretizations of a 1,000 m square heterogeneous unconfined aquifer.  The top and bottom of the aquifer are flat and specified to have elevations of 10 and -30 m, respectively. Models with horizontal discretizations of 200 $\times$ 200, 500 $\times$ 500, 1,000 $\times$ 1,000, 2,000 $\times$ 2,000, and 4,000 $\times$ 4,000 cells per layer were evaluated. Single layer models were evaluated for all horizontal discretizations and multi-layer models were evaluated for select horizontal discretizations to determine the effect that a 7-point stencil, and the resulting additional non-zero entries for each cell, has on GPGPU and parallel CPU execution time. Two- and three-layer models were evaluated for all horizontal discretizations except 4,000 $\times$ 4,000 cells per layer. Ten-layer models were also evaluated for horizontal discretizations ranging from 200 $\times$ 200 to 1,000 $\times$ 1,000 cells per layer. Additional horizontal and vertical discretizations were not evaluated because of GPGPU memory requirements for larger problems. For multi-layer models, a constant layer thickness was used for all but the first layer and was calculated by dividing 30 m by the number of layers. The thickness of layer 1 was calculated to be sum of the top elevation of the model (10 m) and constant thickness used for layers 2 and below.

The heterogeneous hydraulic conductivity distribution (base hydraulic conductivity data) was generated for the 1,000 $\times$ 1,000 cell model (base model) using sequential Gaussian simulation with log-transformed (base 10) parameters based on a mean value of 5 m/d and an exponential variogram having a nugget of 1.023 $\log_{10}((m/d)^{2})$, a variance of 1.23 $\log_{10}((m/d)^{2})$, and length parameter ($a$) of 250.5 m (effective range $\approx$ 750 m). The axis of anisotropy was rotated by 45 degrees to the model grid and a 3.5:1 ratio of horizontal anisotropy was used. For the realization selected and used to evaluate GPGPU and parallel CPU results, the minimum, average, and maximum hydraulic conductivity in the model domain were 0.43, 4.81, and 43.3 m/d, respectively. The distribution of hydraulic conductivity in the model domain is shown in \textbf{figure~\ref{FigModelDomain}}.

@@

%%Tables and Figures
%Table1
\newpage
\begin{table}[hp]
   \caption{Speedup of UPCG solver GPGPU simulations relative to CPU simulations for the same UPCG preconditioner. Speedup values in \textbf{bold} type indicate simulations that did not converge within 50 outer iterations with up to 1,000 inner iterations.}
   \label{TableGPUCPUSpeedup}
   \begin{tabular}{ l | c | c | c |}
 		\cline{2-4}
		& \multicolumn{3}{c|}{Speedup} \\ \cline{1-4}
		\multicolumn{1}{|c|}{Columns $\times$ Rows $\times$ Layers} & Jacobi & MILU0 & GLSPOLY \\ \cline{1-4}
        		\multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 1 } &  1.513 &  0.990 &  2.972 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 1 } &  9.536 &  1.699 & 16.315 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 1 } & 18.871 &  1.404 & 26.786 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 1 } & 24.453 &  1.938 & 33.615 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 4000 $\times$ 4000 $\times$ 1 } & \textbf{26.918} &  1.936 & 35.154 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 2 } &  3.234 &  1.517 &  6.146 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 2 } & 13.420 &  1.742 & 18.665 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 2 } & 19.877 &  1.322 & 25.420 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 2 } & 23.871 &  1.861 & 29.516 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 3 } &  5.234 &  1.517 &  9.231 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 3 } & 15.848 &  1.769 & 20.843 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 3 } & 22.444 &  1.265 & 27.658 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 3 } & 25.171 &  1.870 & 30.760 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 10 } & 11.552 &  1.715 & 17.564 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 10 } & 20.806 &  1.833 & 25.826 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 10 } & 24.981 &  1.889 & 29.888 \\ \cline{1-4}
       \end{tabular}
   \end{table}

%Table2
\newpage
\begin{table}[hp]
   \caption{Speedup of UPCG solver GPGPU simulations relative to simulations performed using the PCG solver with the modified incomplete Cholesky (MIC) preconditioner. CPU times for the PCG solver with the MIC preconditioner are in seconds. Speedup values in \textbf{bold} type indicate simulations that did not converge within 50 outer iterations with up to 1,000 inner iterations.}
   \label{TableGPUSpeedup}
   \begin{tabular}{ l | c | c | c | c |}
 		\cline{2-5}
		& CPU Time & \multicolumn{3}{c|}{Speedup} \\ \cline{1-5}
		\multicolumn{1}{|c|}{Columns $\times$ Rows $\times$ Layers} & MIC & Jacobi & MILU0 & GLSPOLY \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 1 } & 0.6826875 &  0.487 &  1.101 &  0.841 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 1 } & 10.27194 &  1.675 &  1.543 &  2.294 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 1 } & 75.64222 &  3.241 &  1.239 &  2.957 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 1 } & 536.2608 &  1.891 &  1.438 &  2.754 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 4000 $\times$ 4000 $\times$ 1 } & 3891.683 & \textbf{ 1.357} &  1.696 &  2.839 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 2 } & 1.683125 &  1.081 &  1.330 &  1.633 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 2 } & 24.66269 &  3.156 &  1.534 &  3.497 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 2 } & 175.7042 &  4.111 &  1.156 &  3.532 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 2 } & 1058.137 &  1.967 &  1.590 &  2.837 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 3 } & 3.040594 &  1.875 &  1.487 &  2.349 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 3 } & 38.59184 &  3.908 &  1.530 &  4.075 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 3 } & 315.6253 &  5.105 &  1.130 &  4.428 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 3 } & 1962.365 &  2.560 &  1.634 &  3.562 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 10 } & 18.50184 &  5.344 &  1.524 &  4.921 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 10 } & 218.9169 &  8.683 &  1.598 &  7.643 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 10 } & 1621.158 &  9.299 &  1.621 &  8.633 \\ \cline{1-5}
       \end{tabular}
   \end{table}

>>>>

%%% PAGE DIMENSIONS
\usepackage{geometry}
\usepackage{lineno}
\usepackage{color, colortbl}
\usepackage[kernelfbox]{xcolor}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
%%% LANGUAGE HYPHENATION
\usepackage[english]{babel}
%%% HEADERS & FOOTERS
\usepackage{fancyhdr} 
%%% SECTION TITLE APPEARANCE
\usepackage{sectsty}
%%% MATHS RELATED PACKAGES
\usepackage{import}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{setspace}
\usepackage{array}
\usepackage{multirow}
%%% OTHER PACKAGES
\usepackage{booktabs}
\usepackage{array}
\usepackage{paralist}
\usepackage{verbatim}
\usepackage{subfig}
\usepackage{url}

@@

\color{black}Dong and Li (2009) parallelized the MODFLOW PCG solver (Hill 1990) using the OpenMP (OpenMP Architecture Review Board, 2005) programming paradigm and observed speedups as large as 5 on machines with multi-core CPUs. Naff (2008) developed a parallelized linear solver for MODFLOW using the Message Passing Interface (MPI) standard and observed a speedup of 7. In addition to OpenMP and MPI, GPGPUs also \color{blue}present \color{black}an option for linear solver parallelization. In comparison to CPUs, GPGPUs \color{blue}typically have more cores (\textit{e.g.}, 448 cores in the case of the NVIDIA$^{\textregistered}$ Tesla™ C2050 and C2070 GPGPUs), have higher bandwidth, and as a result are capable of executing more floating point operations per second. \color{black}Use of GPGPUs to parallelize linear solvers for MODFLOW is also attractive because of the availability of the CUDA (Compute Unified Device Architecture) application programming interface (API) for NVIDIA GPGPUs, which supports development of parallel GPGPU code in C/C++ or Python, Fortran, Java, and MATLAB using wrappers around the C/C++ implementation. For example, Soni \textit{et al.} (2012) used the CUDA language to develop a GPGPU code to solve computational fluid dynamics problems. For solution of the groundwater flow equation, Ji \textit{et al.} (2012) developed a parallel GPGPU version of MODFLOW-2000 with a Jacobi preconditioner and observed speedups ranging from 1.8 to 3.7.

@@

For GPGPU code implementations, the concept is similar to OpenMP.  The main code is a sequential host code which forks the data into many threads at the invocation of a CUDA \color{blue}functions, called kernels (\textbf{Box 1})\color{black}.  \color{blue}However, as opposed to the few threads spawned by OpenMP, GPGPUs have to potential to spawn thousands to tens-of-thousands of threads to execute the CUDA kernel function (\textbf{Box 2}). The NVIDIA Tesla C2050 used to evaluate the UPCG solver in this study can execute 21,504 concurrent threads. \color{black} With thread counts on the order of tens of thousands, GPGPUs have great potential for parallelization \color{blue} for MODFLOW and other numerical simulators that use Krylov subspace methods to solve large systems of simultaneous linear equations\color{black}.

@@

\subsection*{Problem Description}
GPGPU and parallel CPU results were evaluated using a number of horizontal and vertical model discretizations of a 1,000 m square heterogeneous unconfined aquifer.  The top and bottom of the aquifer are flat and specified to have elevations of 10 and -30 m, respectively. Models with horizontal discretizations of 200 $\times$ 200, 500 $\times$ 500, 1,000 $\times$ 1,000, 2,000 $\times$ 2,000, and 4,000 $\times$ 4,000 cells per layer were evaluated. Single layer models were evaluated for all horizontal discretizations and multi-layer models were evaluated for select horizontal discretizations to determine the effect that a 7-point stencil, and the resulting additional non-zero entries for each cell, has on GPGPU and parallel CPU execution time. Two- and three-layer models were evaluated for all horizontal discretizations except 4,000 $\times$ 4,000 cells per layer. Ten-layer models were also evaluated for horizontal discretizations ranging from 200 $\times$ 200 to 1,000 $\times$ 1,000 cells per layer. Additional horizontal and vertical discretizations were not evaluated because of GPGPU memory requirements for larger problems. For multi-layer models, a constant layer thickness was used for all but the first layer and was calculated by dividing 30 m by the number of layers. \color{blue}Layer 1 also includes the thickness of the aquifer above the arbitrary elevation of 0 m and therefore is 10 m thicker than layer 2.

\color{black}The heterogeneous hydraulic conductivity distribution (base hydraulic conductivity data) was generated for the 1,000 $\times$ 1,000 cell model (base model) using sequential Gaussian simulation with log-transformed (base 10) parameters based on a mean value of 5 m/d and an exponential variogram having a nugget of 1.023 $\log_{10}((m/d)^{2})$, a variance of 1.23 $\log_{10}((m/d)^{2})$, and length parameter ($a$) of 250.5 m (effective range $\approx$ 750 m). The axis of anisotropy was rotated by 45 degrees to the model grid and a 3.5:1 ratio of horizontal anisotropy was used. For the realization selected and used to evaluate GPGPU and parallel CPU results, the minimum, average, and maximum hydraulic conductivity in the model domain were 0.43, 4.81, and 43.3 m/d, respectively. The distribution of hydraulic conductivity in the model domain is shown in \textbf{figure~\ref{FigModelDomain}}.

@@

%%Tables and Figures
%Table1
\newpage
\begin{table}[hp]
   \caption{Speedup of UPCG solver GPGPU simulations relative to CPU simulations for the same UPCG preconditioner. Speedup values in shaded cells indicate simulations that did not converge within 50 outer iterations requiring as much as 1,000 inner iterations.}
   \label{TableGPUCPUSpeedup}
   \begin{tabular}{ l | c | c | c |}
 		\cline{2-4}
		& \multicolumn{3}{c|}{Speedup} \\ \cline{1-4}
		\multicolumn{1}{|c|}{Columns $\times$ Rows $\times$ Layers} & Jacobi & MILU0 & GLSPOLY \\ \cline{1-4}
        		\multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 1 } &  1.513 &  0.990 &  2.972 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 1 } &  9.536 &  1.699 & 16.315 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 1 } & 18.871 &  1.404 & 26.786 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 1 } & 24.453 &  1.938 & 33.615 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 4000 $\times$ 4000 $\times$ 1 } & \cellcolor{gray} 26.918 &  1.936 & 35.154 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 2 } &  3.234 &  1.517 &  6.146 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 2 } & 13.420 &  1.742 & 18.665 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 2 } & 19.877 &  1.322 & 25.420 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 2 } & 23.871 &  1.861 & 29.516 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 3 } &  5.234 &  1.517 &  9.231 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 3 } & 15.848 &  1.769 & 20.843 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 3 } & 22.444 &  1.265 & 27.658 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 3 } & 25.171 &  1.870 & 30.760 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 10 } & 11.552 &  1.715 & 17.564 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 10 } & 20.806 &  1.833 & 25.826 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 10 } & 24.981 &  1.889 & 29.888 \\ \cline{1-4}
       \end{tabular}
   \end{table}

%Table2
\newpage
\begin{table}[hp]
   \caption{Speedup of UPCG solver GPGPU simulations relative to simulations performed using the PCG solver with the modified incomplete Cholesky (MIC) preconditioner. CPU times for the PCG solver with the MIC preconditioner are in seconds. Speedup values in  in shaded cells indicate simulations that did not converge within 50 outer iterations requiring as much as 1,000 inner iterations.}
   \label{TableGPUSpeedup}
   \begin{tabular}{ l | c | c | c | c |}
 		\cline{2-5}
		& CPU Time & \multicolumn{3}{c|}{Speedup} \\ \cline{1-5}
		\multicolumn{1}{|c|}{Columns $\times$ Rows $\times$ Layers} & MIC & Jacobi & MILU0 & GLSPOLY \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 1 } & 0.6826875 &  0.487 &  1.101 &  0.841 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 1 } & 10.27194 &  1.675 &  1.543 &  2.294 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 1 } & 75.64222 &  3.241 &  1.239 &  2.957 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 1 } & 536.2608 &  1.891 &  1.438 &  2.754 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 4000 $\times$ 4000 $\times$ 1 } & 3891.683 & \cellcolor{gray} 1.357 &  1.696 &  2.839 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 2 } & 1.683125 &  1.081 &  1.330 &  1.633 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 2 } & 24.66269 &  3.156 &  1.534 &  3.497 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 2 } & 175.7042 &  4.111 &  1.156 &  3.532 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 2 } & 1058.137 &  1.967 &  1.590 &  2.837 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 3 } & 3.040594 &  1.875 &  1.487 &  2.349 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 3 } & 38.59184 &  3.908 &  1.530 &  4.075 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 3 } & 315.6253 &  5.105 &  1.130 &  4.428 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 3 } & 1962.365 &  2.560 &  1.634 &  3.562 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 10 } & 18.50184 &  5.344 &  1.524 &  4.921 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 10 } & 218.9169 &  8.683 &  1.598 &  7.643 \\ \cline{1-5}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 10 } & 1621.158 &  9.299 &  1.621 &  8.633 \\ \cline{1-5}
       \end{tabular}
   \end{table}

%Table3
\newpage
\begin{table}[hp]
   \caption{Speedup of UPCG solver OpenMP simulations, with 4, 7, and 14 threads, relative to simulations performed using the PCG solver with the modified incomplete Cholesky (MIC) preconditioner. Speedup values in in shaded cells indicate simulations that did not converge within 50 outer iterations requiring as much as 1,000 inner iterations.}
   \label{TableOpenMPSpeedup}
   \begin{tabular}{ l | c | c | c |}
 		\cline{2-4}
		& \multicolumn{3}{c|}{Speedup -- 4 / 7 /14 Threads} \\ \cline{1-4}
		\multicolumn{1}{|c|}{Columns $\times$ Rows $\times$ Layers} & Jacobi & MILU0 & GLSPOLY \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 1 } & 1.166/0.614/0.614 &  2.446/1.511/1.423 & 1.185/0.779/1.241 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 1 } & 0.447/0.332/0.357 & 1.637/1.197/1.164 & 0.571/0.364/0.535 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 1 } & 0.409/0.419/0.359 & 1.488/1.463/1.194 & 0.313/0.323/0.309 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 1 } & 0.185/0.182/0.151 & 1.691/1.622/1.286 & 0.225/0.272/0.198 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 4000 $\times$ 4000 $\times$ 1 } & \cellcolor{gray}0.111/0.099/0.108 & 1.619/1.523/1.670 &  0.222/0.241/0.245 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 2 } & 1.113/1.090/1.004 & 1.954/2.258/1.567 & 1.171/1.147/1.626 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 2 } &  0.536/0.567/0.460 & 1.500/1.521/1.151 & 0.525/0.625/0.556 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 2 } & 0.527/0.472/0.411 & 1.543/1.474/1.227 & 0.396/0.295/0.335 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 2 } & 0.211/0.198/0.168 & 1.401/1.693/1.418 & 0.237/0.313/0.244 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 3 } & 0.872/1.067/0.897 & 1.433/1.901/1.300 & 0.827/1.080/1.346 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 3 } &  0.577/0.482/0.527 & 1.444/1.417/1.328 & 0.548/0.579/0.598 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 3 } &  0.593/0.562/0.479 & 1.519/1.590/1.377 & 0.389/0.340/0.385 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 3 } &  0.245/0.246/0.215 & 1.746/1.478/1.544 & 0.313/0.286/0.295 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 10 } & 1.066/1.059/1.215 & 1.503/1.574/1.426 & 0.858/0.899/1.186 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 10 } & 0.970/1.014/1.019 & 1.563/1.604/1.400 & 0.780/0.799/0.913 \\ \cline{1-4}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 10 } & 0.835/0.968/0.928 & 1.720/1.734/1.651 & 0.815/0.798/0.914 \\ \cline{1-4}
       \end{tabular}
   \end{table}

%Table4

\newpage
\begin{table}[hp]
\caption{Convergence summary for simulations using the PCG solver with the MIC preconditioner, the GMG solver with the ILU0 preconditioner, and the UPCG solver with the Jacobi, MILU0, and GLSPOLY preconditioners. The total number of outer and inner iterations are shown (outer/inner). Shaded cells indicate simulations that did not converge within 50 outer iterations requiring as much as 1,000 inner iterations.}
   \label{tab:UPCGITSummary}
   \begin{tabular}{ l | c | c | c | c | c |}
 		\cline{2-6}
		& \multicolumn{5}{c|}{Convergence History} \\ \cline{1-6}
		\multicolumn{1}{|c|}{Columns $\times$ Rows $\times$ Layers} & MIC & GMG & Jacobi & MILU0 & GLSPOLY \\ \cline{1-6}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 1 } & 6/140 & 5/7 & 5/955 & 6/136 & 6/130 \\ \cline{1-6}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 1 } & 7/312 & 5/9 & 6/2629 & 7/298 & 7/352 \\ \cline{1-6}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 1 } & 9/574 & 6/10 & 8/4624 & 8/547 & 7/765 \\ \cline{1-6}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 1 } & 8/1022 & 7/14 & 21/17862 & 8/980 & 8/1752 \\ \cline{1-6}
        \multicolumn{1}{| l |}{ 4000 $\times$ 4000 $\times$ 1 } & 9/1859 & 8/16 & \cellcolor{gray}50/50000 & 9/1758 & 9/3257 \\ \cline{1-6}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 2 } & 6/151 & 5/7 & 5/811 & 6/145 & 5/123 \\ \cline{1-6}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 2 } & 6/334 & 7/12 & 5/2073 & 6/315 & 5/276 \\ \cline{1-6}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 2 } & 7/597 & 8/16 & 6/4102 & 7/562 & 6/633 \\ \cline{1-6}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 2 } & 7/902 & 20/33 & 17/15342 & 7/864 & 5/1363 \\ \cline{1-6}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 3 } & 6/167 & 5/6 & 5/788 & 5/154 & 5/124 \\ \cline{1-6}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 3 } & 5/335 & 6/11 & 5/1967 & 5/319 & 5/261 \\ \cline{1-6}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 3 } & 7/685 & 7/14 & 6/4059 & 6/633 & 5/637 \\ \cline{1-6}
        \multicolumn{1}{| l |}{ 2000 $\times$ 2000 $\times$ 3 } & 6/1069 & 23/33 & 16/14848 & 6/1003 & 5/1359 \\ \cline{1-6}
        \multicolumn{1}{| l |}{ 200 $\times$ 200 $\times$ 10 } & 6/291 & 5/7 & 5/955 & 6/273 & 5/164 \\ \cline{1-6}
        \multicolumn{1}{| l |}{ 500 $\times$ 500 $\times$ 10 } & 5/546 & 6/9 & 5/1908 & 5/512 & 5/275 \\ \cline{1-6}
        \multicolumn{1}{| l |}{ 1000 $\times$ 1000 $\times$ 10 } & 7/1014 & 7/13 & 6/3913 & 6/955 & 5/521 \\ \cline{1-6}
       \end{tabular}
   \end{table}

%Figure 1
\newpage
